---
title: "HELIUS depression microbiome; Bos et al; Nature communications"
output:
  html_document:
    toc: yes
    toc_depth: 1
    number_sections: yes
    theme: united
  pdf_document:
    toc: yes
    toc_depth: '1'
---

<style>
p.caption {
  font-size: 0.8em;
}
</style>

```{r setup rmd, include=FALSE}
# setting global options for the rmd markdown
knitr::opts_chunk$set(echo = T) # include chunck code
knitr::opts_chunk$set(cache = T) # use chuck results from cache if not modified (speed)
```

This page will hold the code for some of the bioinformatics and statistical methods used for the article [The gut microbiota and depressive symptoms across ethnic groups](https://www.nature.com/articles/s41467-022-34504-1)


# ASV inference 

# Statistical analysis

## Generic SPSS code Linear regression Models 

```{r, echo=F}
flextable::flextable(rio::import("analysis_code.xlsx"))
```

## Generic SPSS code Linear regression Models 1 - 3

REGRESSION

  /MISSING LISTWISE
  
  /STATISTICS COEFF OUTS CI(95) R ANOVA CHANGE ZPP
  
  /CRITERIA=PIN(.05) POUT(.10)
  
  /NOORIGIN 
  
  /DEPENDENT [Dependent variable]
  
  /METHOD=ENTER [Predictor variable]
  
  /METHOD=ENTER [Covariate list Model 1]
  
  /METHOD=ENTER [Covariate list Model 2]   
  
  /METHOD=ENTER [Covariate list Model 3]
  
  /SCATTERPLOT=([Dependent variable] ,*ZRESID)
  
  /RESIDUALS HISTOGRAM(ZRESID) NORMPROB(ZRESID)




## Generic SPSS code ANOVA (testing interaction ethnicity by alpha diverity or relative abundcances)

UNIANOVA H1_PHQ9_sumscore BY H1_EtnTotaal H1_geslacht WITH shannon H1_lft

  /METHOD=SSTYPE(3)
  
  /INTERCEPT=INCLUDE
  
  /PRINT ETASQ DESCRIPTIVE PARAMETER
  
  /CRITERIA=ALPHA(.05)
  
  /DESIGN=H1_EtnTotaal*shannon H1_EtnTotaal shannon H1_geslacht H1_lft




## SPSS code Linear regression forward selection PcoA


REGRESSION

  /MISSING LISTWISE
  
  /STATISTICS COEFF OUTS CI(95) R ANOVA CHANGE ZPP
  
  /CRITERIA=PIN(.05) POUT(.10)
  
  /NOORIGIN 
  
  /DEPENDENT H1_PHQ9_sumscore
  
  /METHOD=FORWARD
  

PC01BC PC02BC PC03BC PC04BC PC05BC PC06BC PC07BC PC08BC PC09BC PC010BC PC011BC PC012BC 
    PC013BC PC014BC PC015BC PC016BC PC017BC PC018BC PC019BC PC020BC .  

Retained: 	Weighted UniFrac 	PC02 PC03 PC07 PC011 PC014 PC019

Bray-Curtis 		PC02BC PC03BC PC04BC PC05BC PC014BC PC016BC PC018BC



## Generic SPSS code Linear regression Models Beta-Diversity

REGRESSION

  /MISSING LISTWISE
  
  /STATISTICS COEFF OUTS CI(95) R ANOVA CHANGE ZPP
  
  /CRITERIA=PIN(.05) POUT(.10)
  
  /NOORIGIN 
  
  /DEPENDENT [Dependent variable]
  
  /METHOD=ENTER [Ethicity  or  Empty]
  
  /METHOD=ENTER [Covariate list Model 1 -3]
  
  /METHOD=ENTER [retained principal components Beta-diversity]
  
  /SCATTERPLOT=([Dependent variable] ,*ZRESID)
  
  /RESIDUALS HISTOGRAM(ZRESID) NORMPROB(ZRESID)
